Auditory Scene Similarity Retrieval and Classification with Relevance-based Quantization of Scattering Features

### 2016-12-20 paper/structureScene.tex
edit Results log compression
Fix standard deviation
more on the point conclusion

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
Consequently, few well-designed datasets are readily available for evaluation purposes.

@@

Logarithmic compression provides a small boost to acoustic scene classification over all but one fold of the DCASE 2013 private dataset.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale. This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).

@@

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification. In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and more diverse applications, such as pleasantness prediction \cite{lafaySoundscapePilot} as well as other emerging tasks in computational bioacoustics.

>>>>

This works falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause, wimmer2013sampling}. Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
Consequently, few well-designed datasets are readily available for evaluation purposes.

@@

On the DCASE 2013 dataset, logarithmic compression provides a small, albeit consistent boost to acoustic scene classification.
Indeed, as discussed in subsection \ref{sec:logcomp}, and observed in Figure \ref{fig:compression}, the statistical assumption of normality for scattering coefficients is only approximately satisfied once mapped to a logarithmic, decibel-like scale. This transformation improves the discriminative power of the SVMs trained on scattering coefficients, regardless of the chosen temporal integration strategy (early or late) and the inner product kernel (linear or Gaussian RBF).

@@

This paper presents a new approach for modeling acoustic scenes which utilizes scattering transforms at small scales and a cluster-based representation at large scales. Compared to traditional models based on BOF and summary statistics, this representation allows for the characterization of distinct sound events superimposed on a stationary texture, a concept which has strong grounding in the cognitive psychology literature. We demonstrate the usefulness of scattering-based representations for auditory scenes and the value of late integration using experiments in auditory scene classification. In order to extend the selectivity potential of late integration using SVMs to the unsupervised case, we develop the cluster-based model and validate it using experiments on acoustic scene similarity retrieval. For this task, we show significant improvements over the traditional BOF and summary statistics models based on both standard MFCCs and scattering features.
These outcomes shall be studied further in future work by considering larger databases and emerging tasks in computational bioacoustics \cite{wimmer2013sampling}.
